\documentclass[unicode,11pt,a4paper,oneside,numbers=endperiod,openany]{scrartcl}

\usepackage{assignment}
\usepackage{textcomp}
\usepackage{url}
\hyphenation{PageRank}
\hyphenation{PageRanks}

\definecolor{Brown}{cmyk}{0,0.81,1,0.60}
\definecolor{Green}{rgb}{0.516, 0.5976, 0.480}
\definecolor{OliveGreen}{cmyk}{0.64,0,0.95,0.40}
\definecolor{CadetBlue}{rgb}{0.17,0.30,0.547}
\definecolor{Red}{rgb}{0.684,0.156,0.32}
\lstset{
  language=matlab,
  basicstyle=\ttfamily,
  frame=ltrb,
  framesep=5pt,
  basicstyle=\scriptsize,
  stringstyle=\ttfamily\color{Brown}\bfseries,
  commentstyle=\ttfamily\color{Green}\bfseries,
  keywordstyle=\ttfamily\color{CadetBlue}\bfseries,
  identifierstyle=\ttfamily, 
  tabsize=2,
  showstringspaces=true,
  numbers=none
}



\begin{document}

\setassignment
\setduedate{Thursday 12 November 2015, 10:30}
\serieheader{Introduction to PDEs (LAB)}{Academic Year 2015/2016}{Instructor: Prof. Rolf Krause}{Dr. Drosos Kourounis}{Assignment 7 - Iterative Krylov-based methods}{}

\section*{Solution of Poisson's equation using iterative Krylov methods}
Solve the same equation you solved in LAB6 replacing the backslash
with each one of the methods provided in the MATLAB folder of our repository. 
Enforcing Dirichlet boundary conditions renders the system non-symmetric.
For the non-symmetric version of the system you should use Krylov-space methods
designed for non-symmetric systems of equations, nameley CGS, BICGSTAB and GMRES.
These can be used with both the incomplete Cholesky (ichol) and incomplete LU
(ilu) preconditioners. Report iterations and residual for each choice of method
and preconditioner in a table where the horizontal row is $N=50,100,200$,
and the vertical row of the table is the Krylov method. In order for this to work
as expected, you need to write a function that returns the application of the preconditioner
on a vector, and instead of providing the preconditioner matrix $M$ on each one of these
methods you should provide a pointer to function. Read MATLAB documentation to understand
how to do this.

\begin{table}[h]
\centering
\begin{tabular}{|l|r|r|r|}
\hline
                  &  \multicolumn{3}{|c|}{$N$} \\
\hline
 Krylov method    & 50  & 100 & 200 \\
\hline
GMRES(ilu)        & iters (residual)    &  iters (residual)   &  iters (residual)\\
GMRES(ichol)      &     &     &  \\
\hline
CGS(ilu)          &     &     &  \\
CGS(ichol)        &     &     &  \\
\hline
BICGSTAB(ilu)     &     &     &  \\
BICGSTAB(ichol)   &     &     &  \\
\hline
\end{tabular}
\caption{Performance of non-symmetric Krylov methods with ilu0 and ichol preconditioners on a grid of size $N \times N \times N$.}
\end{table}
Implement the symmetric version of Dirichlet conditions. Now you can use ichol
with conjugate gradient CG and QMR as well. Create another table 
You should try both the preconditioners provided from MATLAB 
\begin{table}[h]
\centering
\begin{tabular}{|l|r|r|r|}
\hline
                  &  \multicolumn{3}{|c|}{$N$} \\
\hline
 Krylov method    & 50  & 100 & 200 \\
\hline
CG                &  iters (residual)   &  iters (residual)   &  iters (residual)\\
\hline
QMR               &     &     &  \\
\hline
GMRES             &     &     &  \\
\hline
CGS               &     &     &  \\
\hline
BICGSTAB          &     &     &  \\
\hline
\end{tabular}
\caption{Performance of symmetric and non-symmetric Krylov methods with ichol 
         preconditioners with Dirichlet conditions preserving the symmetry of 
         the original stiffness matrix, on a grid of size $N \times N \times N$.}
\end{table}
Plot the log of the residual norm (y-axis) as a function of the iterations (x-axis)
for all the methods of Table 2, for two cases. One plot for $N=50$ and one for $N=200$ side-by-side.
Use different colors for each method and make sure you add a legend in your plots. 


\section*{Using FEniCS}
Solve the same problem using FEniCS. Choose the Krylov solvers you see below and the hypre\_amg
or petsc\_amg and icc preconditioners. Create a similar table with Table 2. Finally make a table
of the running times for both FEniCS and your code.
\begin{table}[h]
\centering
\begin{tabular}{|l|r|r|r|}
\hline
                  &  \multicolumn{3}{|c|}{$N$} \\
\hline
 Krylov method    & 50  & 100 & 200 \\
\hline
CG (HYPRE)        &                     &                     & 3 (1.067e-8)  \\
CG (KSP)          &                     &                     & 7 (3.008e-8)  \\
CG (ICC)          &                     &                     &68 (4.739e-7)  \\
\hline
GMRES (HYPRE)     &                     &                     & 3 (1.016e-4) \\
GMRES (KSP)       &                     &                     & 7 (1.016e-4) \\
GMRES (ICC)       &                     &                     & 60 (2.2e-4) \\
\hline
BICGSTAB (HYPRE)  &                     &                     & 3 (1.512e-8) \\
BICGSTAB (KSP)    &                     &                     & 7 (7.126e-8) \\
BICGSTAB (ICC)    &                     &                     & 48 (1.769e-7) \\
\hline
\end{tabular}
\caption{Performance of symmetric and non-symmetric Krylov methods with icc, petsc\_amg and hypre\_amg  preconditioners with Dirichlet conditions preserving the symmetry of 
         the original stiffness matrix, on a grid of size $N \times N \times N$.}
\end{table}


\end{document}
