\documentclass[unicode,11pt,a4paper,oneside,numbers=endperiod,openany]{scrartcl}

\usepackage{assignment}
\usepackage{textcomp}
\usepackage{url}
\hyphenation{PageRank}
\hyphenation{PageRanks}

\definecolor{Brown}{cmyk}{0,0.81,1,0.60}
\definecolor{Green}{rgb}{0.516, 0.5976, 0.480}
\definecolor{OliveGreen}{cmyk}{0.64,0,0.95,0.40}
\definecolor{CadetBlue}{rgb}{0.17,0.30,0.547}
\definecolor{Red}{rgb}{0.684,0.156,0.32}
\lstset{
  language=matlab,
  basicstyle=\ttfamily,
  frame=ltrb,
  framesep=5pt,
  basicstyle=\scriptsize,
  stringstyle=\ttfamily\color{Brown}\bfseries,
  commentstyle=\ttfamily\color{Green}\bfseries,
  keywordstyle=\ttfamily\color{CadetBlue}\bfseries,
  identifierstyle=\ttfamily, 
  tabsize=2,
  showstringspaces=true,
  numbers=none
}



\begin{document}

\setassignment
\setduedate{Thursday 15 October 2015, 10:30}
\serieheader{Introduction to PDEs (LAB)}{Academic Year 2015/2016}{Instructor: Prof. Rolf Krause}{TA: Drosos Kourounis}{Assignment 3 - Finite Element Solution of Poisson's equation in 2D}{}


\section*{Solution of Poisson's equation}
We seek the discrete solution of Poisson's equation
\begin{align*}
-\nabla^2 u(x, y) &= f(x,y), \; \in \Omega \\
u                 &= u_0(x,y), \; \in \partial \Omega
\end{align*}
where $\Omega = [0,1]^2$.
\begin{enumerate}
\item Derive the analytical expression of $f(x,y)$ so that the exact 
solution of the PDE is \[u_0(x,y) =10 x + \tanh(10x - 10)\] 
\item Write {\small MATLAB} code that solves the PDE 
implementing the steps following\footnote{ 
You should include in your submission only the implementations of the
functions marked with red and not the whole \small{MATLAB} code.}
\end{enumerate}

\section{Mesh generation}
Write a \small{MATLAB} function that generates a $N_x \times N_y$ quadrilateral grid.
The input arguments of the function should be \lstinline+N_x, N_y, grid_type+. If third argument \lstinline+grid_type+
is equal to \lstinline+'quadrilaterals'+ then you do not split each quadrilateral
to 2 triangles as you did in the previous Assignment. 
The output of the function should be the matrix of the connectivity of the grid 
(\lstinline+Elements+) and the matrix of the coordinates of the grid points (\lstinline+Points+). 
The matrix \lstinline+Elements+ should by of size $N_e \times N_\text{v}$,
where the $N_e$ number of elements and $N_\text{v}$ the number of vertices. The matrix
\lstinline+Points+, should be of size $N \times d$, where $N$ the number of points and $d$ the number of 
space dimensions.
\begin{lstlisting}[numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=10pt, identifierstyle=\ttfamily\color{Red}\bfseries]
function [mesh] = makeGrid(L_x, L_y, N_x, N_y, grid_type)
\end{lstlisting}
The output \lstinline+mesh+ is a MATLAB structure that contains all the information that are needed for the mesh generation
and mesh description, namely, \lstinline+delta_x, delta_y, N_x, N_y, L_x, L_y, N_e, N_v, N, Elements, Points, PointMarkers+. The parameters \lstinline+L_x,L_y+ represent the length and height of the domain at each direction. Here both are equal to $1$, since our domain $\Omega = [0, 1]^2$. However, it does not hurt to write the code for the general case. 

The array \lstinline+PointMarkers+ should be 0 at every point except those points that belong to the boundary where they should be 1. We will 
need it later to enforce Dirichlet boundary conditions to our Discrete system of equations.
Use the MATLAB function \lstinline+writeMeshAsVTKFile.m+ in the \lstinline+MATLAB+ folder
and call it passing the structure mesh and the name of the output file.
This will dump the grid as a VTK file that you can visualize using visualization package ViSiT. The declarations follow.
\begin{lstlisting}[numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=10pt]
function writeMeshAsVTKFile(mesh, vtkfile)
\end{lstlisting}



\section{Finite element assembly}
Use the following code sample for the assembly of your discrete operators.
\begin{lstlisting}[numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=10pt]
function [M,K,b] = assembleDiscreteOperators(mesh)
  N   = mesh.N;
  Ne  = mesh.N_e;
  M   = zeros(N,N); K = zeros(N,N);
  b   = zeros(N,1);
  for e=1:N_e
    Me = makeMe(e, mesh);
    Ke = makeKe(e, mesh);
    fp = makebe(e, mesh);
    I = mesh.Elements(e, :);
    M(I, I) = M(I, I) + Me;
    K(I, I) = K(I, I) + Ke;
    b(I)    = b(I) + Me*fp;
  end % e loop
end
\end{lstlisting}
Provide implementations of the individual 
functions needed for forming the element mass matrix \lstinline+M_e+, element Laplacian \lstinline+K_e+ 
and the element rhs \lstinline+b_e+. Use the provided MATLAB routines \lstinline+Ksymbolic.m, Msymbolic.m+ that can be found in the \lstinline+MATLAB+ folder. These routines will compute the mass and stiffness matrices only for the particular grid type. The result will be wrong in the general case where we need to use quadrature.
\begin{lstlisting}[numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=10pt, identifierstyle=\ttfamily\color{Red}\bfseries] 
function Me = makeMe(e, mesh);
function Ke = makeKe(e, mesh);
function fp = makebe(e, mesh);
\end{lstlisting}
Keep in mind that \lstinline+f_p+ should hold the values of the function $f(x,y)$
evaluated at the nodes of the $e$th element,  for example
$f_p = \left [ f(x^e_1, y^e_1) \; f(x^e_2, y^e_2) \; f(x^e_3, y^e_3)  \; f(x^e_4, y^e_4)\right ]^T$.


\section{Solution}
Provide a function that solves the linear system $K u_h = b$, and use it to obtain the solution of the linear system
you constructed in the previous step. Provide a function that enforces Dirichlet boundary conditions for all the points that have been marked with 1. For the particular exercise you can stop the loop at the first point of the grid, since we want the solution to satisfy Neumann conditions on the boundary. Dirichlet conditions will destroy the symmetry of the matrix. Can you modify the system such that the symmetry is preserved without screwing up the solution?  
\begin{lstlisting}[numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=10pt, identifierstyle=\ttfamily\color{Red}\bfseries] 
function x = solve(K, b)
\end{lstlisting}

\section{Convergence study}
Compare the solution you obtained with the exact one in the first part $u_0(x,y)$ by computing the 
$L^2$ and $H^1$ norms of the error and plot them using logarithmic scale on both axes
as a function of $h = \max{(\delta x, \delta y)}$ for $N_x = N_y = 5,10,20,40,80,160$.
Note that the discrete $L^2$ and $H^1$ norms are easilly obtained from
\begin{align*}
\|u - u_h\|_{L^2} &= \sqrt{(u-u_h)^T M (u-u_h)} \\
\|u - u_h\|_{H^1} &= \sqrt{(u-u_h)^T M (u-u_h) + (u-u_h)^T K (u-u_h)},
\end{align*}
where $u_h$ is the vector of the discrete finite element solution you 
obtained by solving the linear system $K u_h = b$ and $u$ is the vector
of the exact values of the solution. The exact solution
is given by $u_0(x,y)$ evaluated at the grid points. Delive a single $\log\log$ plot where you show the 
reduction of the error with respect to the mesh size $h$ for both the triangular mesh and the quadrilateral mesh.
You are free to use sparse matrices to speed up the solution of the related linear systems.


\section{Visualization}
Use ViSiT to visualize the solution for each different mesh resolution. 
\end{document}
